### 2024-01-26 parts/appendix.tex
preparing the witdrawal

    Remark that if $(v_1, \dots, v_k) \equiv (w_1, \dots, w_k)$ then
    $\pi(\mu(v_1), \dots, \mu(v_k)) = \pi(\mu(w_1), \dots, \mu(w_k))$ because
    $\mu(u_i^\omega) = \mu(u_i)^\omega$ is an idempotent element of the monoid
    $M$. In particular, $f(w)$ can be computed as the sum over equivalence
    classes for $\equiv$, of the value of $\pi$ on one representative,
    which belongs to $\Nat$.

    Therefore, it suffices to prove that the number of partitions of $w$ in a
    single equivalence class for $\equiv$ is a polynomial $P(X_1, \dots, X_p)$
    with \kl{positive coefficients} for its \kl{maximal monomials}.

@@

\begin{proof}[Proof of 
    \cref{derivation-monomials:fact}
    ]
    Let $M \defined X_1^{k_1} \cdots X_n^{k_n}$,
    where $k_1, \dots, k_n > 0$ and $n > 0$.
    Let us compute
    \begin{align*}
        \Diff{K}{M} &= \prod_{i = 1}^n X_i^{k_i} - \prod_{i = 1}^n (X_i - K)^{k_i} \\
                    &= \sum_{p = 1}^n \frac{K}{X_p - K} \prod_{i = 1}^n (X_i - K)^{k_i} 
    \end{align*}
    Therefore, 
    the \kl{maximal monomials} of $\Diff{K}{M}$
    are precisely
    of the form $K \frac{\prod_{i = 1}^n X_i^{k_i}}{X_p}$ for some $1 \leq p \leq i$,
    which \kl{strictly divide} $M$.
    Let $T$ be a \kl{monomial} that \kl{strictly divides} $M$,
    then, $T = \prod_{i = 1}^n X_i^{t_i}$, 
    for all $1 \leq i \leq n$, $t_i \leq k_i$,
    and there exists $1 \leq p \leq n$ such that$ t_j < k_j$.
    In particular, $T$ \kl{divides}
    $K \frac{\prod_{i = 1}^n X_i^{k_i}}{X_p}$.
\end{proof}

\begin{proof}[Proof of \cref{discrete-deriv-linear:fact}]
    TODO or to skip.
\end{proof}

\begin{proof}[Proof of \cref{maxi-monomials-submodular:fact}]
\end{proof}

\begin{proof}[Proof of \cref{derivation-monomials:fact}]
\end{proof}

\begin{proof}[Proof of \cref{derivation-simplifies:lemma}]
    Let $P \in \Rel[\vec{X}]$ and
    $K \in \Nat$.
    Leveraging 
    \cref{derivation-monomials:fact,discrete-deriv-linear:fact,maxi-monomials-submodular:fact}:
    \begin{align*}
        \MaximalMonomials\left(\Diff{K}{P}\right)
        &=
        \MaximalMonomials\left(\sum_{M \in \Monomials(P)} \Diff{K}{M}\right) \\
        &\hoareleq
        \bigcup_{M \in \Monomials(P)} 
        \MaximalMonomials(\Diff{K}{M}) \\
        &\hoarele
        \MaximalMonomials(P)
        \quad .
        \qedhere
    \end{align*}
\end{proof}
\begin{proof}[Proof of \cref{derivation-stabilises-correct:lem}]
    If $P$ is constant, then $P_2 = 0$, and $\Diff{K}{P_1} = 0$,
    hence for $K = 0$, $Q \in \CorrectPoly$.
    Otherwise, let $\nu \colon \vec{X} \topartial \Nat$,
    and let us analyze the \kl{maximal monomials} $T$
    of $\restr{Q}{\nu}$.
    Because $T$ is a \kl{monomial} of $\restr{Q}{\nu}$,
    it \kl{divides} some \kl{monomial} in $Q$, and
    using \cref{derivation-monomials:fact},
    it \kl{divides} some \kl{monomial} $M$ in 
    $\Diff{K}{P_1}$.
    Now, the monomial $T$ is computed as follows:
    \begin{align*}
        T &= 
        \sum_{S \in \Monomials(Q), T \divides S}
        \restr{S}{\nu} \\
          &= 
          \sum_{S \in \Monomials(\Diff{K}{P_1}), T \divides S}
            \restr{S}{\nu}
        + 
          \sum_{S \in \Monomials(P_2), T \divides S}
            \restr{S}{\nu}
        \\
          &\geq
          \sum_{S \in \Monomials(\Diff{K}{P_1}), T \divides S}
            \restr{S}{\nu}
          - 
            \left|
          \sum_{S \in \Monomials(P_2), T \divides S}
            \restr{S}{\nu}
            \right|
    \end{align*}

\end{proof}

\begin{proof}[Proof of
    \cref{decidability-correct:lem}]
    Let $P \in \Rel[\vec{X}]$. We prove by induction on
    $\MaximalMonomials(P)$ that whether $P \in \CorrectPoly$
    is decidable.

    If $P$ is a constant polynomial, then
    $P \in \CorrectPoly$ if and only if $P = n \in \Nat$
    which is decidable.

    Otherwise, $P$ is not a constant polynomial, and let us write $P = P_1 +
    P_2$ where $P_1$ is the sum of the \kl{maximal monomials} of $P$. If $P_1
    \not \in \Nat[\vec{X}]$, then $P \not \in \CorrectPoly$, and this is
    decidable. Hence, we can assume that $P_1 \in \Nat[\vec{X}]$. Using
    \cref{derivation-stabilises-correct:lem}, there exists a computable bound
    $K$, such that if $P \in \CorrectPoly$, then t $Q \defined (\Diff{K}{P_1} +
    P_2) \in \CorrectPoly$. Leveraging, \cref{derivation-simplifies:lemma}, we
    also know that $\MaximalMonomials(Q) \hoarele \MaximalMonomials(P)$, and
    therefore that we can decide whether $Q \in \CorrectPoly$. Notice that if
    $Q \not \in \CorrectPoly$, then $P \not \in \CorrectPoly$, hence we can
    assume that $Q \in \CorrectPoly$.

    Similarly, for partial valuations
    $\nu \colon \vec{X} \topartial \set{0, \dots, K}$
    fixing at least one indeterminate, $\MaximalMonomials(\restr{P}{\nu})
    \hoarele \MaximalMonomials{P}$. 
    We can check for all these partial valuations that
    $\restr{P}{\nu} \in \CorrectPoly$ by induction hypothesis,
    and if one test is not satisfied, then $P \not \in \CorrectPoly$.

    Let us now conclude that if all the above checks answered correctly,
    then $P \in \CorrectPoly$. Let $\nu \colon \vec{X} \topartial \Nat$
    be a partial valuation. Then, 
    one of the three following cases arise:
    \begin{itemize}
        \item 
            Either,
        $\nu$ fixes \emph{some} indeterminate $X_i$ to a value $n_i < K$.
        In that case, let us define $\delta$ as the partial function that coincides
        with $\nu$ on its domain
        $\dom(\delta) = \dom(\nu) \setminus \set{ X_i }$.
        Then, $\restr{P}{\nu} = \restr{\restr{P}{X_i = n_i}}{\delta}$.
        As we previously checked that $\restr{P}{X_i = n_i} \in \CorrectPoly$,
        we conclude in particular that 
        $\restr{P}{\nu}$ has \kl{non-negative} \kl{maximal monomials}.

        \item Or, for all $X_i$ in the (potentially empty) domain of $\nu$, $\nu(X_i)
            \geq K$. In that case, let us write $\delta \defined
            \translate{-K}(\nu)$, that is the partial map $\delta \colon
            \vec{X} \topartial \Nat$ having the same domain as $\nu$, and such
            that for all $X_i \in \dom(\nu)$, $\delta(X_i) \defined \nu(X_i) -
            K$. Recall that $Q = \Diff{K}{P_1} + P_2$, and therefore that
            $\translate{K}(P) = P_1 + \translate{K}(Q)$. By linearity,
            $\restr{\translate{K}(P)}{\delta} = \restr{P_1}{\delta} +
            \restr{\translate{K}(Q)}{\delta}$, and we conclude that the
            \kl{maximal monomials} of $\restr{\translate{K}(P)}{\delta}$ are
            \kl{non-negative}.
            Since $\restr{\translate{K}(P)}{\delta} = \translate{K}{\restr{P}{\nu}}$,
            and since
            $\translate{K}$ leaves \kl{maximal monomials} invariant,
            we conclude that
            the \kl{maximal monomials} of $\restr{P}{\nu}$
            are the same as those of $\restr{\translate{K}(P)}{\delta}$,
            and are therefore \kl{non-negative}.
            \qedhere
    \end{itemize}
\end{proof}

@@

    To prove \cref{d-t-transl:item} $\implies$ \cref{d-t-correct:item}, let us
    remark that if $P$ is \kl{non-negative} and $\translate{K}(P) \in
    \Nat[\vec{X}]$, then $\translate{K}(P) \in \CorrectPoly$ thanks to
    \cref{n-poly-n-poly:example}. Then, for every partial valuation $\nu \colon
    \vec{X} \topartial \Nat$, $\translate{K}(\restr{P}{\nu}) \in
    \Nat[\vec{X}]$, and in particular conclude that \kl{maximal monomials} of
    $\translate{K}(\restr{P}{\nu})$ are exactly the \kl{maximal monomials} of
    $\restr{P}{\nu}$ thanks to \cref{translation-maximal:fact}, and are therefore
    \kl{non-negative}.

@@

    Let us prove that for all partial functions $\nu \colon \vec{X} \topartial
    \Nat$, the translation $\translate{L_m + K}(\restr{P}{\nu})$ belongs to
    $\Nat[\vec{X}]$. Without loss of generality, $\nu$ fixes only
    indeterminates that appear in $P$. If $\nu$ fixes at least one
    indeterminate $X_i$ to a value $\nu(X_i) = k \leq K$, then we can write
    $\nu = \mu [X_i = k]$, where $\mu$ does not fix the value of $X_i$. By
    construction, we know that $\translate{L_i}(\restr{\restr{P}{X_i =
    k}}{\mu})$ belongs to $\Nat[\vec{X}]$, and therefore that $\translate{K +
    L_m}(\restr{P}{\nu})$ does so too. The only case left is when all the
    variables appearing in $\nu$ are greater than $K$.
    In that case, $\nu = \translate{K}(\mu)$ for some partial 
    valuation $\mu$.
    In this case,
    $\translate{K}(\restr{P}{\translate{K}(\mu)}
    = \restr{\translate{K}(P)}{\mu}$
    thanks to 
    \cref{discrete-deriv-linear:fact}.
    Remark that by definition,
    $\translate{K}(P) = P_1 + Q$,
    hence that
    $\translate{K + L_m}(P) = \translate{L_m}(P_1) + \translate{L_m}(Q)$
    belongs to $\Nat[\vec{X}]$.
    Finally,
    we conclude that $\restr{\translate{K+L_m}(P)}{\mu}$
    belongs to $\Nat[\vec{X}]$
    because
    the latter is closed under partial applications.
\end{proof}

>>>>

    Remark that if $(v_1, \dots, v_k) \equiv (w_1, \dots, w_k)$ then their
    production are the same, i.e., $\pi(\mu(v_1), \dots, \mu(v_k)) =
    \pi(\mu(w_1), \dots, \mu(w_k))$. This holds because $\mu(u_i^\omega) =
    \mu(u_i)^\omega$ is an idempotent element of the monoid $M$. In particular,
    $f(w)$ can be computed as the sum over equivalence classes for $\equiv$, of
    the value of $\pi$ on one representative, which belongs to $\Nat$.

    Therefore, it suffices to prove that the number of partitions of $w$ in a
    single equivalence class for $\equiv$ is a polynomial $P(X_1, \dots, X_p)$
    with \kl{non-negative} \kl{maximal monomials}.

@@

\begin{proof}[Proof of 
    \cref{derivation-monomials:fact}
    ]
    Let $M \defined X_1^{k_1} \cdots X_n^{k_n}$,
    where $k_1, \dots, k_n > 0$ and $n > 0$.
    Let us compute
    \begin{align*}
        \Diff{K}{M} &= \prod_{i = 1}^n X_i^{k_i} - \prod_{i = 1}^n (X_i - K)^{k_i} \\
                    &= \sum_{p = 1}^n \frac{K}{X_p - K} \prod_{i = 1}^n (X_i - K)^{k_i} 
    \end{align*}
    Therefore, 
    the \kl{maximal monomials} of $\Diff{K}{M}$
    are precisely
    of the form $K \frac{\prod_{i = 1}^n X_i^{k_i}}{X_p}$ for some $1 \leq p \leq i$,
    which \kl{strictly divides} $M$.
    Let $T$ be a \kl{monomial} that \kl{strictly divides} $M$,
    then, $T = \prod_{i = 1}^n X_i^{t_i}$, 
    for all $1 \leq i \leq n$, $t_i \leq k_i$,
    and there exists $1 \leq p \leq n$ such that$ t_j < k_j$.
    In particular, $T$ \kl{divides}
    $K \frac{\prod_{i = 1}^n X_i^{k_i}}{X_p}$.
\end{proof}

\begin{proof}[Proof of \cref{derivation-simplifies:lemma}]
    Let $P \in \Rel[\vec{X}]$ and
    $K \in \Nat$.
    Leveraging 
    \cref{derivation-monomials:fact,discrete-deriv-linear:fact,maxi-monomials-submodular:fact}:
    \begin{align*}
        \MaximalMonomials\left(\Diff{K}{P}\right)
        &=
        \MaximalMonomials\left(\sum_{M \in \Monomials(P)} \Diff{K}{M}\right) \\
        &\hoareleq
        \bigcup_{M \in \Monomials(P)} 
        \MaximalMonomials(\Diff{K}{M}) \\
        &\hoarele
        \MaximalMonomials(P)
        \quad .
        \qedhere
    \end{align*}
\end{proof}

\begin{proof}[Proof of \cref{derivation-stabilises-correct:lem}]
    If $P$ is constant, then $P_2 = 0$, and $\Diff{K}{P_1} = 0$,
    hence for $K = 0$, $Q \in \CorrectPoly$.

    Otherwise, let $\alpha$ be the maximal absolute value 
    of coefficients appearing in $P$,
    and $D$ be the maximal number of divisors for
    a monomial in $P$.
    We define $K \defined \alpha \times D + 1$.

    Let $\nu \colon \vec{X} \topartial \Nat$, we are going to prove that
    $\restr{Q}{\nu}$ has \kl{non-negative} \kl{maximal monomials}. To that end,
    let us consider $T$ such a maximal monomial, and let us write $T = \beta
    T_1$ so that $T_1$ has coefficient $1$ and $\beta \in \Rel$. Because we
    will need to normalize other monomials in this way, we define for the rest
    of the proof the function $\mathsf{unit}$ that maps a monomial to the
    monomial obtained by replacing its constant term by $1$ if the monomial is
    non-zero, and leaving the monomial unchanged otherwise.

    Let us write $\mathcal{D}_1$ to be the set of \kl{monomials} of
    $\Diff{K}{P_1}$ that are \kl{divided} by $T$. Similarly, let us write
    $\mathcal{D}_2$ for the set of monomials of $\translate{K}(P_2)$ that are
    \kl{divided} by $T$. Thanks to \cref{derivation-monomials:fact}, we know
    that every element of $\mathcal{D}_2$ divides some element in
    $\mathcal{D}_1$. We write $\mathcal{D}_2^M$ for the set of monomials $S \in
    \mathcal{D}_2$ such that $S$ \kl{divides} $M$.

    Now, remark that the coefficient $\beta$ is obtained as follows:
    \begin{align*}
        \beta &= \sum_{M \in \mathcal{D}_1} \restr{M / T_1}{\nu}
               + \sum_{M \in \mathcal{D}_2} \restr{M / T_1}{\nu} \\
              &\geq \sum_{M \in \mathcal{D}_1} \restr{M / T_1}{\nu}
              - \sum_{M \in \mathcal{D}_2} \left|\restr{M / T_1}{\nu}\right|
              \\
              &\geq
              \sum_{M \in \mathcal{D}_1}
              \left[
              \restr{M / T_1}{\nu}
              -
              \sum_{S \in \mathcal{D}_2^M}
              \left|\restr{S/T_1}{\nu}\right|
              \right]
              \\
              &\geq
              \sum_{M \in \mathcal{D}_1}
              \restr{\mathsf{unit}(M)/T_1}{\nu}
              \left[
              K
              - \alpha D
              \right]
              \\
              &\geq 0 \quad . \qedhere
    \end{align*}
\end{proof}

@@

    We will first prove the equivalence between \cref{d-t-transl:item} and
    \cref{d-t-correct:item} without keeping track of the constant $K$, and then
    notice that this $K$ is actually computable. This will simplify the reading
    process.

    We will first prove that \cref{d-t-transl:item} $\implies$
    \cref{d-t-correct:item}, for any choice of $K \in \Nat$. Indeed, let us
    consider some partial valuation $\nu \colon \vec{X} \topartial \Nat$. By
    assumption, $\translate{K}(\restr{P}{\nu}) \in \Nat[\vec{X}]$, and in
    particular conclude that \kl{maximal monomials} of
    $\translate{K}(\restr{P}{\nu})$ are exactly the \kl{maximal monomials} of
    $\restr{P}{\nu}$ thanks to \cref{translation-maximal:fact}, and are
    therefore \kl{non-negative}.

@@

    Let us prove that for all partial functions $\nu \colon \vec{X} \topartial
    \Nat$, the translation $\translate{L_m + K}(\restr{P}{\nu})$ belongs to
    $\Nat[\vec{X}]$. Without loss of generality, $\nu$ fixes only
    indeterminates that appear in $P$. If $\nu$ fixes at least one
    indeterminate $X_i$ to a value $\nu(X_i) = k \leq K$, then we can write
    $\nu = \mu [X_i = k]$, where $\mu$ does not fix the value of $X_i$. By
    construction, we know that $\translate{L_i}(\restr{\restr{P}{X_i =
    k}}{\mu})$ belongs to $\Nat[\vec{X}]$, and therefore that $\translate{K +
    L_m}(\restr{P}{\nu})$ does so too. The only case left is when all the
    variables appearing in $\nu$ are greater than $K$.
    In that case, $\nu = \translate{K}(\mu)$ for some partial 
    valuation $\mu$.
    In this case,
    $\translate{K}(\restr{P}{\translate{K}(\mu)}
    = \restr{\translate{K}(P)}{\mu}$
    thanks to 
    \cref{discrete-deriv-linear:fact}.
    Remark that by definition,
    $\translate{K}(P) = P_1 + Q$,
    hence that
    $\translate{K + L_m}(P) = \translate{L_m}(P_1) + \translate{L_m}(Q)$
    belongs to $\Nat[\vec{X}]$.
    Finally,
    we conclude that $\restr{\translate{K+L_m}(P)}{\mu}$
    belongs to $\Nat[\vec{X}]$
    because
    the latter is closed under partial applications.

    Now, remark that in the proof, we have computed $K$ recursively from $P$,
    and only applied partial functions using values below $K$. This not only
    shows that $K$ is computable from $P$, but also that
    \cref{d-t-transl-fin:item} and \cref{d-t-transl:item} are equivalent.
\end{proof}

### 2024-01-26 parts/correction.tex
preparing the witdrawal

\begin{lemma}
    \label{n-poly-combinatorics:lem}
    Let $f$ be a  \kl{$\Nat$-polyregular} function. 
    There exists $\omega \in \Nat$
    such that for all $p \in \Nat$,
    for all $\alpha_0, \dots, \alpha_p \in \Sigma^*$,
    for all $u_1, \dots, u_p \in \Sigma^*$,
    there exists a polynomial $P \in \Rel[X_1, \dots, X_p]$
    whose \kl{maximal monomials} are \kl{non-negative},
    and such that for all $X_1, \dots, X_p \geq \omega$:
    \begin{equation*}
        f\left(
            \alpha_0 \prod_{i = 1}^p u_i^{\omega X_i} \alpha_i
        \right)
        = P(X_1, \dots, X_p) \quad .
    \end{equation*}
\end{lemma}

@@

The following \cref{derivation-stabilises-correct:lem} is the main
combinatorial argument of this section. Note that we must be careful
about it. \textbf{TODO: write a nice thing here!}.

@@

    \textbf{Induction:} Assume that $P$ is not a constant polynomial, and let
    us write $P = P_1 + P_2$ where $P_1$ is the sum of the \kl{maximal
    monomials} of $P$. We compute a bound $K$ such that $Q \defined
    (\Diff{K}{P_1} + P_2) \in \CorrectPoly$ using
    \cref{derivation-stabilises-correct:lem}. Thanks to
    \cref{derivation-simplifies:lemma}, we also know that $\MaximalMonomials(Q)
    \hoarele \MaximalMonomials(P)$. By induction hypothesis, $Q$ is
    \kl{represented} by a \kl{star-free $\Nat$-polyregular function}
    which is effectively computable.

    Let us now remark that $P_1 \in \Nat[\vec{X}]$, and is therefore
    (effectively) \kl{represented} by a \kl{star-free $\Nat$-polyregular function}
    (\cref{n-poly-n-poly:example}).
    Furthermore, $\translate{K}(Q)$ is also (effectively)
    \kl{represented} by a \kl{star-free $\Nat$-polyregular function}
    (see \cref{translation-invariance:fact}).
    As a consequence, $\translate{K}(P)$ is (effectively) \kl{represented} 
    by a function $f_\Delta$,
    since it is obtained as the following sum:
    \begin{equation*}
        \translate{K}(P) 
        = P_1 + \translate{K}(Q) \quad .
    \end{equation*}

@@

While \cref{lem:correct-to-n-rat} provides effective conversion, it does not
explicitly state that the membership is decidable to keep the proof clearer. A
similar proof scheme can be followed to conclude that membership is decidable.

\begin{lemma}
    \label{decidability-correct:lem}
    It is decidable whether a polynomial 
    $P \in \Rel[\vec{X}]$
    belongs to $\CorrectPoly$.
\end{lemma}

@@

Let us also state that polynomials in $\CorrectPoly$ are, up to suitable
translations, polynomials in $\Nat[\vec{X}]$. Beware that partial applications
are still needed in this characterization, as \cref{bad-poly-translate:ex}
illustrates.

\begin{lemma}
    \label{derivation-translation:lem}
    Let $P \in \Rel[\vec{X}]$, the following are equivalent:
    \begin{enumerate}
        \item \label{d-t-correct:item} $P \in \CorrectPoly$,
        \item \label{d-t-transl:item}
            There exists $K \in \Nat$,
            such that for every partial
            function $\nu \colon \vec{X} \topartial \Nat$,
            $\translate{K}(\restr{P}{\nu}) \in \Nat[\vec{X}]$.
    \end{enumerate}
\end{lemma}

>>>>

\begin{lemma}
    \label{n-poly-combinatorics:lem}
    Let $f$ be a  \kl{$\Nat$-polyregular} function. 
    There exists a computable $\omega \in \Nat$
    such that for all $p \in \Nat$,
    for all $\alpha_0, \dots, \alpha_p \in \Sigma^*$,
    for all $u_1, \dots, u_p \in \Sigma^*$,
    there exists a polynomial $P \in \Rel[X_1, \dots, X_p]$
    whose \kl{maximal monomials} are \kl{non-negative},
    and such that for all $X_1, \dots, X_p \geq \omega$:
    \begin{equation*}
        f\left(
            \alpha_0 \prod_{i = 1}^p u_i^{\omega X_i} \alpha_i
        \right)
        = P(X_1, \dots, X_p) \quad .
    \end{equation*}
\end{lemma}

@@

The following \cref{derivation-stabilises-correct:lem} is the main
combinatorial argument of this section. It leverages the positivity of the
\kl{maximal monomials} to compute a shift large enough to make lower degree
coefficients positive. This lemma will be lifted to a given polynomial $P$ by
noticing that if $K \in \Nat$, and $P = P_1 + P_2$, then $\translate{K}(P) =
P_1 + \Diff{K}{P_1} + \translate{K}(P_2)$.

@@

    \textbf{Induction:} Assume that $P$ is not a constant polynomial, and let
    us write $P = P_1 + P_2$ where $P_1$ is the sum of the \kl{maximal
    monomials} of $P$. We compute a bound $K$ such that $Q \defined
    (\Diff{K}{P_1} + \translate{K}(P_2)) \in \CorrectPoly$ using
    \cref{derivation-stabilises-correct:lem}. Thanks to
    \cref{derivation-simplifies:lemma}, we also know that $\MaximalMonomials(Q)
    \hoarele \MaximalMonomials(P)$. By induction hypothesis, $Q$ is
    \kl{represented} by a \kl{star-free $\Nat$-polyregular function} which is
    effectively computable.

    Let us now remark that $P_1 \in \Nat[\vec{X}]$, and is therefore
    (effectively) \kl{represented} by a \kl{star-free $\Nat$-polyregular
    function} (\cref{n-poly-n-poly:example}). As a consequence,
    $\translate{K}(P) = P_1 + Q$ is (effectively) \kl{represented} by a
    function $f_\Delta$.

@@

While \cref{lem:correct-to-n-rat} provides effective conversion, it does not
explicitly state that the membership is decidable to keep the proof clearer. A
similar proof scheme can be followed to conclude that membership is decidable,
and even show that elements in $\CorrectPoly$ are, up to suitable translations,
polynomials in $\Nat[\vec{X}]$. Beware that partial applications are still
needed in this characterization, as \cref{bad-poly-translate:ex} illustrates.

\begin{lemma}
    \label{derivation-translation:lem}
    Let $P \in \Rel[\vec{X}]$, 
    there exists a computable $K \in \Nat$
    such that the following are equivalent:
    \begin{enumerate}
        \item \label{d-t-correct:item} $P \in \CorrectPoly$,
        \item \label{d-t-transl:item}
            for 
            all $\nu \colon \vec{X} \topartial \Nat$,
            $\translate{K}(\restr{P}{\nu}) \in \Nat[\vec{X}]$,
        \item \label{d-t-transl-fin:item}
            for all
            $\nu \colon \vec{X} \topartial \set{0, \dots, K}$,
            $\translate{K}(\restr{P}{\nu}) \in \Nat[\vec{X}]$.
    \end{enumerate}
\end{lemma}

\begin{corollary}
    \label{decidability-correct:lem}
    One can decide membership in $\CorrectPoly$.
\end{corollary}
\begin{proof}
    Given $P \in \Rel[\vec{X}]$,
    one applies \cref{derivation-translation:lem}
    to obtain a bound $K$, and then 
    computes $\translate{K}(\restr{P}{\nu})$ for all 
    partial valuations $\nu \colon \vec{X} \topartial \set{0, \dots, K}$.
    Then $P \in \CorrectPoly$ if and only if all those polynomials
    belong to $\Nat[\vec{X}]$, which is decidable.
\end{proof}

@@

### 2024-01-26 parts/prelim.tex
preparing the witdrawal

\begin{lemma}[TODO]
    \label{polynomial-rational-polyreg:fact}
    Let $f \in \NRat$ (resp. $f \in \ZRat$), then
    the following are equivalent:
    \begin{enumerate}
        \item $f \in \NPoly$ (resp. $\ZPoly$);
        \item $f$ is a \kl{polyregular function} with output
            in $\set{1}^*$,
            postcomposed with $\polysum$
            (resp. with output in $\set{-1,+1}^*$);
        \item $f$ has \kl{polynomial growth};
        \item $f$ can be constructed
            using $+$, $\cauchy$, and multiplication by
            an integer $m \in \Nat$ (resp. $m \in \Rel$)
            starting from
            indicator functions $\ind{L}$ of regular languages;
        \item $f$ belongs to
            $\Span{\mathbb{S}}(\setof{\vcount{\varphi}}{\varphi(\vec{x}) \in \MSO})$
            with $\mathbb{S} = \Nat$
            (resp. $\mathbb{S} = \Rel$).
    \end{enumerate}
\end{lemma}

@@

\AP Inside $\ZPoly$, there is an infinite hierarchy of classes, defined using
the \kl{growth rate}. Namely, one defines $\ZPoly[k]$ as the functions $f \in
\ZPoly$ such that $\growth{f}(n) = \bigO(n^k)$, i.e., have \kl{growth rate}
$k$. Similarly, we let $\NPoly[k] = \NPoly \cap \ZPoly[k]$.

>>>>

\begin{lemma}[TODO]
    \label{polynomial-rational-polyreg:fact}
    Let $f \in \NRat$ (resp. $f \in \ZRat$), then
    the following are equivalent:
    \begin{enumerate}
        \item $f \in \NPoly$ (resp. $\ZPoly$);
        \item $f$ is a \kl{polyregular function} with output
            in $\set{1}^*$,
            postcomposed with $\polysum$
            (resp. with output in $\set{-1,+1}^*$);
        \item $f$ has \kl{polynomial growth};
        \item $f$ can be constructed
            using $+$, $\cauchy$, and multiplication by
            an integer $m \in \Nat$ (resp. $m \in \Rel$)
            starting from
            indicator functions $\ind{L}$ of regular languages;
        \item \label{npoly-counting-mso:item} $f$ belongs to
            $\Span{\mathbb{S}}(\setof{\vcount{\varphi}}{\varphi(\vec{x}) \in \MSO})$
            with $\mathbb{S} = \Nat$
            (resp. $\mathbb{S} = \Rel$).
    \end{enumerate}
\end{lemma}

@@

\textbf{This is wrong, we should talk about the number of variables,
    and say that we do not know in the case of $\NSF$ if it corresponds
    to the minimal number of variables needed!}
\AP Inside $\ZPoly$, there is an infinite hierarchy of classes, defined using
the \kl{growth rate}. Namely, one defines $\ZPoly[k]$ as the functions $f \in
\ZPoly$ such that $\growth{f}(n) = \bigO(n^k)$, i.e., have \kl{growth rate}
$k$. Similarly, we let $\NPoly[k] = \NPoly \cap \ZPoly[k]$.
